Evidence of widspread maintenence of functional composition in vertebrate communities

### 2021-09-07 paper/manuscript.Rmd
add citation
Minor changes

While functional loss is frequently cited as one of the most pressing concerns of the anthropocene [@cardinale2012; @dirzo2014; @young2016], it is not necessarily inevitable even in scenarios of species loss [@daz2001]. Forecasts of functional loss range from negligible [@gallagher2013] to dire [@petchey2002; @pimiento2020]. And while some observed trends show significant functional loss [@flynn2009] others document no loss even in some of the most heavily impacted communities [@edwards2013; @matuoka2020].
It is therefore critical to establish whether or not functional loss is the general rule for communities, and at what rate functional loss is occurring.

@@

Our final dataset included 2,443 timeseries from 53 studies in 21 countries and 15 biomes and 13 different traits (Fig 1).
Only four studies (consisting of 11 timeseries) came from Amphibian studies due to the limited availability of amphibian timeseries and low species richness values for assemblages (Table 1).
For a full breakdown of studies and their characteristics, see the supplement.

@@

<!-- the majority of studies (86%) experienced a significant temporal change in species richness and all functional diversity metrics. 92% of studies experienced a significant trend for species richness, 88% for functional richness, 98% for functional divergence, 96% for functional evenness, and 30% for Jaccard similarity. Only 4 (cite studies) out of 53 studies experienced no significant trend in species richness and were also characterized by significantly decreasing functional evenness. There was only one significant pairwise correlation between the metrics, a positive relationship between Functional Divergence and Functional Evenness, indicating that communities experiencing an increase in Divergence frequently experience an increase in Evenness. -->

@@

Communities demonstrated almost universal maintenance of functional composition.
While the majority of the studies (\~70%) included in our data experienced significant species turnover, only three experienced a significant shift in any functional dimension.
This suggests certain characteristics of the functional space are conserved even in the face of significant change in species identity, namely the size of the functional space occupied by the community (FRic) and the distribution of species and individuals within that space (FEve and FDiv).
On average, species additions have similar functional characteristics as lost species and therefore maintain the functional space structure.
<!-- language to distinguish between "functional traits" and a species' role in the context of the functional space -->

>>>>

While functional loss is frequently cited as one of the most pressing concerns of the anthropocene [@cardinale2012; @dirzo2014; @young2016], it is not necessarily inevitable even in scenarios of species loss [@daz2001].
Forecasts of functional loss range from negligible [@gallagher2013] to dire [@petchey2002; @pimiento2020].
And while some observed trends show significant functional loss [@flynn2009] others document no loss even in some of the most heavily impacted communities [@edwards2013; @matuoka2020].
On paleoecological time scales, functional composition is often conserved alongside species diversity even in the face of drastic environmental change [@jackson2015; @barnosky2005].
It is therefore critical to establish whether or not functional loss is the general rule for communities on time scales relevant for current environmental impacts.

@@

Our final dataset included 2,443 timeseries from 53 studies in 21 countries and 15 biomes and 13 different traits (Fig 1).
The earliest sample was in 1923 and the most recent was in 2016.
Only four studies (consisting of 11 timeseries) came from Amphibian studies due to the limited availability of amphibian timeseries and low species richness values for assemblages (Table 1).
For a full breakdown of studies and their characteristics, see the supplement.

@@

@@

Communities demonstrated almost universal maintenance of functional composition.
While the majority of the studies (\~70%) included in our data experienced significant species turnover, only three experienced a significant shift in any functional dimension.
This suggests certain characteristics of the functional space are conserved even in the face of significant change in species identity, specifically the size of the functional space occupied by the community (FRic) and the distribution of species and individuals within that space (FEve and FDiv).
On average, species additions have similar functional characteristics as lost species and therefore maintain the functional space structure.
<!-- While we did not explicitly assess how turnover rates compare to background turnover, --> <!-- language to distinguish between "functional traits" and a species' role in the context of the functional space -->
